---
title: "Basic Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{a01_basic_usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(timeseriesdb)
```


## Starting Point

This guide assumes you have installed the {timeseriesdb} R package and a local PostgreSQL client library successfully. We also assume that you have a PostgreSQL database running somewhere that you can connect to. (*For the remainder of this guide we assume PostgreSQL runs on a local docker container accessible via port 1111.*)

## General Note

In {timeseriesdb}, functions to interact with the database use the `db_` prefix. 
Basic read and store functions are an exception to this rule as they should be easy to find
front and center. As opposed to that more complex, less common operations rather 
ask for organization in a group functions.

## Create a Database Connection

No matter whether you want to read or write, you'll need a database connection. {timeseriesdb}
offers a simple wrapper function around {DBI}'s more universal `dbConnect`:

```{r, eval=FALSE}
con <- db_connection_create(
  dbname = "postgres",
  user = "dev_admin",
  host = "localhost",
  passwd = "dev_admin",
  port = 1111
)
```

The resulting PostgreSQL connection object can be passed on to all of {timeseriesdb}'s
database interaction functions. 


## Store Time Series 

Let's take two time series from the {datasets} R package that ships with R (monthly AirPassengers, JohnsonJohnson quarterly earnings per share) and put them into a named list. The {timeseriesdb} R package uses lists of time series or long format time series `data.tables` to store one or more time series into PostgreSQL. 

```{r, eval=FALSE}
tsl <- list(
  AirPassengers = AirPassengers,
  JohnsonJohnson = JohnsonJohnson
)
db_ts_store(con, tsl, "main",
  schema = "tsdb_test"
)
```
 
## Read Time Series
 
Now let's read the series we have just stored to the database into R again. 
Note that `db_ts_read` returns a list of time series whose elements' names
are the time series keys (unique identifiers) from the database. In this case
example we use our [{dygraphs}](https://cran.r-project.org/package=dygraphs) package to visualize both series. 
 
```{r, eval=FALSE}

out <- db_ts_read(con, c(
  "JohnsonJohnson",
  "AirPassengers"
),
schema = "tsdb_test"
)

library(dygraphs)
library(xts)
xts_for_viz <- do.call("cbind",lapply(out,as.xts))
dygraph(xts_for_viz) %>% 
  dySeries("JohnsonJohnson", axis="y2")
```

```{r,echo=FALSE,message=FALSE, warning=FALSE}
out <- list(
  AirPassengers = AirPassengers,
  JohnsonJohnson = JohnsonJohnson
)
library(dygraphs)
library(xts)
xts_for_viz <- do.call("cbind",lapply(out,as.xts))
dygraph(xts_for_viz) %>% 
  dySeries("JohnsonJohnson", axis="y2")
```



## Basic Metadata 

In {timeseriesdb}, the most basic way to store data descriptions is to assign non-translated (unlocalized) meta information at the time series level. In order to store meta information
to the database, lets first create a `tsmeta` object in R. 


```{r}
library(timeseriesdb)
tsm <- create_tsmeta(
  JohnsonJohnson = list(
    title = "Historical Quarterly earnings (dollars) per Johnson & Johnson share",
    source = "Shumway, R. H. and Stoffer, D. S. (2000) Time Series Analysis and its Applications. Second Edition. Springer. Example 1.1"
  ),
  AirPassengers = list(
    title = "The classic Box & Jenkins airline data. Monthly totals of international airline passengers",
    source = "Box, G. E. P., Jenkins, G. M. and Reinsel, G. C. (1976) Time Series Analysis, Forecasting and Control. Third Edition. Holden-Day. Series G."
  )
)

tsm
```

Now, pass on the `tsmeta` object to the database. 


```{r, eval=FALSE}

db_store_ts_metadata(con,
  tsm,
  valid_from = Sys.Date(),
  schema = "tsdb_test"
)
```
Note that the default locale for the write operation is NULL. 
This means there is only one language and data descriptions are 
not translated to different languages. If you want to learn how to add localized meta information
or meta information at the vintage or dataset level, please read the [advanced usage vignette](a02_advanced_usage.html). 

Reading the data from the database is just as easy: 


```{r, eval=FALSE}
tsm <- db_read_ts_metadata(con, c(
  "JohnsonJohnson",
  "AirPassengers"
),
schema = "tsdb_test"
)
```



## Advanced Use

{timeseriesdb} offers a plethora of features beyond mere storage of time series and 
their descriptions: {timeseriesdb} can keep track of different versions of the 
same time series (vintages), group time series into datasets, store user specific 
collection of series, manage access on time series level and manage release dates. 

The learn more about these advanced features read the [advanced usage vignette](a02_advanced_usage.html).








