---
title: "Admin Usage"
subtitle: "Administrator Documentation"
author: "Matthias Bannert and Severin Thöni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Starting Point

This guide assumes you have installed the {timeseriesdb} R package and a local PostgreSQL client library successfully. We also assume that you have a PostgreSQL database running somewhere that you can connect to. (*For the remainder of this guide we assume PostgreSQL runs on a local docker container accessible via port 1111.*)

## General Note

With {timeseriesdb}, access rights can be customized quite a bit, so roles will 
be quite specific to the institution that uses it. Still though, {timeseriesdb}
ships with a basic idea of access levels and roles. This document focuses on 
administrator usage as opposed to users, i.e., researchers that want to work
with the time series in the database. 


## Datasets

In {timeseriesdb}, all time series have to be part of a dataset. If the user 
stores a time series w/o specifying a dataset that series is assigned to the 'default'
dataset. Datasets allow to group data and store data descriptions at the 
dataset level. Also, datasets come in handy for admin operations like 
deletion of an entire dataset. 

### Registering a New Dataset

```{r, eval=FALSE}
library(kofdata)
arr <- get_time_series("ch.zrh_airport.arrival.total")
dep <- get_time_series("ch.zrh_airport.departure.total")

store_time_series(con, c(arr,dep), schema = "tsdb_test")

db_create_dataset(con, "ch.zrh_airport",
                  set_description = "Arrivals and Departures ZH Airport",
                  schema = "tsdb_test")

# see, the new dataset is there... 
db_list_datasets(con, schema = "tsdb_test")

```
Apparently, the dataset is empty. We need to assign some time series first.
Make sure the time series you want to register to a set are actually around.
Unlike `store_time_series` assigning series to a dataset is *not* a
vintage level operation, it applies to all versions of a series. This is why time series are assigned to the 'default' schema at first when they are stored and can be moved later on. 

```{r, eval=FALSE}
db_assign_dataset(con,
                  ts_keys = c("ch.zrh_airport.arrival.total",
                              "ch.zrh_airport.departure.total"),
                  set_name = "ch.zrh_airport",
                  schema = "tsdb_test")

```

**Note** that there is no such thing as remove time series from dataset. Either
you delete a time series entirely or you assign it to another dataset.
Think of a registry similar to a software registry. 

### DANGER ZONE: Deleting Datasets



## Adding New Access Levels
