---
title: "Admin Usage"
subtitle: "Administrator Documentation"
author: "Matthias Bannert and Severin Thöni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Starting Point

This guide assumes you have installed the {timeseriesdb} R package and a local PostgreSQL client library successfully. We also assume you're connect to your PostgreSQL database and are familiar with the basic ideas of {timeseriesdb}. (*For the remainder of this guide we assume PostgreSQL runs on a local docker container accessible via port 1111.*)

## General Note

With {timeseriesdb}, access rights can be customized quite a bit, so in the end 
roles will be quite specific to any institution that uses it. Still though, {timeseriesdb}
ships with a basic idea of access levels and roles. This document focuses on 
administrator usage as opposed to users, i.e., researchers that want to work
with the time series in the database. 


## Datasets

In {timeseriesdb}, all time series have to be part of a dataset. If the user 
stores a time series w/o specifying a dataset that series will be assigned to the 'default'
dataset. Datasets allow to group data and store data descriptions at the 
dataset level. Also, datasets come in handy for admin operations like 
deletion of an entire dataset. Therefore new datasets can only be added by admins while
time series can be assigned to existing datasets by ordinary users with write access.

### Registering a New Dataset

```{r, eval=FALSE}
library(kofdata)
arr <- get_time_series("ch.zrh_airport.arrival.total")
dep <- get_time_series("ch.zrh_airport.departure.total")

store_time_series(con, c(arr,dep), schema = "tsdb_test")

db_create_dataset(con, "ch.zrh_airport",
                  set_description = "Arrivals and Departures ZH Airport",
                  schema = "tsdb_test")

# see, the new dataset is there... 
db_list_datasets(con, schema = "tsdb_test")

```
Apparently, the dataset is empty. We need to assign some time series first.
Make sure the time series you want to register to a set are actually around.
Unlike `store_time_series` assigning series to a dataset is *not* a
vintage level operation, it applies to all versions of a series. This is why time series are assigned to the 'default' schema at first when they are stored and can be moved later on. 

```{r, eval=FALSE}
db_assign_dataset(con,
                  ts_keys = c("ch.zrh_airport.arrival.total",
                              "ch.zrh_airport.departure.total"),
                  set_name = "ch.zrh_airport",
                  schema = "tsdb_test")

```

**Note** that there is no such thing as remove time series from dataset. Either
you delete a time series entirely or you assign it to another dataset.
Think of a registry similar to a software registry. 


### DANGER ZONE: Deleting Datasets

Admin users can delete an entire dataset. The power to do so comes with 
great responsibility as it will delete not only the dataset registration entry
but ALL time series, their versions and all meta information associated to the
dataset AND the time series. A deleted dataset CANNOT by restored. The only way
to get a deleted dataset back is a PostgreSQL level backup outside of {timeseriesdb}.

```{r, eval=FALSE}
db_dataset_delete(con, "dataset_name", schema = "tsdb_test")
```

Note, datasets can only be deleted interactively as you have to confirm the 
deletion of a dataset by typing its name to the console. 


## Release Management

A new feature introduced with {timeseriesdb} 1.0 is its support for release management. 
In official and economic statistics, data are usually updated periodically. Often 
data are released on release date known in advance. Data may be available inside an institution but 
only be released later on. In order to facilitate the publication process. 

```{r, eval=FALSE}
db_create_release(con,
                  id = "jul_2020_air",
                  datasets = "ch.zrh_airport",
                  title = "Airport Data from Zurich Airport",
                  release_date = "2020-07-08",
                  target_year = 2020,
                  target_period = 7,
                  target_frequency = 12,
                  note = "Just a test release",
                  schema = "tsdb_test"
                  )
```

In {timeseriesdb}, releases have an id which uniquely identifies the release.
We recommend to choose a short human readable text identifier. Releases relate to a dataset and therefore are assigned at the dataset level. This is a good reason to make use of the dataset feature as opposed to storing all series in the default set. Releases have a target year and target period because the release date itself may not safely identify the period it relates to (assume July data get released at the beginning of August). 

See also .... 


## Access Levels & Roles

{timeseriesdb} can control access at the time series and even at the version level. 
The basic installation ships with three access levels: public, main and restricted. 
The idea of the public level is that every users with access to the {timeseriesdb}
database schema can read public series. The idea of the 'main' level is to give 
all users inside an organization access to the assigned data. 'restricted' is
designed to limit access to subgroups within an organization. We guess you see 
where this is going. By adding further access levels, admins can add more narrow
user groups for parts of the data.

```{r, eval=FALSE}

```








