---
title: "Admin Usage"
subtitle: "Administrator Documentation"
author: "Matthias Bannert and Severin Thöni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Admin Usage}
  %\VignetteEngine{rmarkdown::render}
  \usepackage[utf8]{inputenc}
---

## Starting Point

This guide assumes you have installed the {timeseriesdb} R package and a local PostgreSQL client library successfully. We also assume you're connect to your PostgreSQL database and are familiar with the basic ideas of {timeseriesdb}. (*For the remainder of this guide we assume PostgreSQL runs on a local docker container accessible via port 1111.*)

## General Note

With {timeseriesdb}, access rights can be customized quite a bit, so in the end 
roles will be quite specific to any institution that uses it. Still though, {timeseriesdb}
ships with a basic idea of access levels and roles. This document focuses on 
administrator usage as opposed to users, i.e., researchers that want to work
with the time series in the database. 


## Datasets

In {timeseriesdb}, all time series have to be part of a dataset. If the user 
stores a time series w/o specifying a dataset that series will be assigned to the 'default'
dataset. Datasets allow to group data and store data descriptions at the 
dataset level. Also, datasets come in handy for admin operations like 
deletion of an entire dataset. Therefore new datasets can only be added by admins while
time series can be assigned to existing datasets by ordinary users with write access.

### Registering a New Dataset

```{r, eval=FALSE}
library(kofdata)
arr <- get_time_series("ch.zrh_airport.arrival.total")
dep <- get_time_series("ch.zrh_airport.departure.total")

db_ts_store(con, c(arr,dep), schema = "tsdb_test")

db_dataset_create(con, "ch.zrh_airport",
                  set_description = "Arrivals and Departures ZH Airport",
                  schema = "tsdb_test")

# see, the new dataset is there... 
db_dataset_list(con, schema = "tsdb_test")

```

Apparently, the dataset is empty. We need to assign some time series first.



### DANGER ZONE: Deleting Datasets

Admin users can delete an entire dataset. The power to do so comes with 
great responsibility as it will delete not only the dataset registration entry
but ALL time series, their versions and all meta information associated to the
dataset AND the time series. A deleted dataset CANNOT by restored. The only way
to get a deleted dataset back is a PostgreSQL level backup outside of {timeseriesdb}.

```{r, eval=FALSE}
db_dataset_delete(con, "dataset_name", schema = "tsdb_test")
```

Note, datasets can only be deleted interactively as you have to confirm the 
deletion of a dataset by typing its name to the console. 


### Trimming History

{timeseriesdb} stores every new update of a registered time series as a new version of this series. 
That way it keeps track of all changes and revision. For unrevised series and/or series that are updated very frequently, this behavior may be over the top and admins may want to delete versions that are older than 
a particular threshold. {timeseriesdb} offer [db_dataset_trim_history()](../reference/db_dataset_trim_history.html) and [db_ts_trim_history()](../reference/db_ts_trim_history.html) to facilitate this type of maintenance work. In production setups this functions might be used in batch mode / workflow automation. 


## Release Management

A new feature introduced with {timeseriesdb} 1.0 is its support for release management. 
In official and economic statistics, data are usually updated periodically. Often 
data are released on release date known in advance. Data may be available inside an institution but 
only be released later on. In order to facilitate the publication process. 

```{r, eval=FALSE}
db_release_create(con,
                  id = "jul_2020_air",
                  datasets = "ch.zrh_airport",
                  title = "Airport Data from Zurich Airport",
                  release_date = "2020-07-08",
                  target_year = 2020,
                  target_period = 7,
                  target_frequency = 12,
                  note = "Just a test release",
                  schema = "tsdb_test"
                  )
```

In {timeseriesdb}, releases have an id which uniquely identifies the release.
We recommend to choose a short human readable text identifier. Releases relate to a dataset and therefore are assigned at the dataset level. This is a good reason to make use of the dataset feature as opposed to storing all series in the default set. Releases have a target year and target period because the release date itself may not safely identify the period it relates to (assume July data get released at the beginning of August). 

[db_release_update()](../reference/db_release_update.html) updates a release date and [db_release_cancel(db_release_cancel.html)] lets an admin 
cancel a release. Past releases cannot be cancelled. Note that, there are several 
release related functions that are not restricted to admins and are therefore 
discussed in the function reference and the user vignettes. 


## Access 

The access concept of {timeseriesdb} relies on revoking all rights to query the
data directly and allows users instead to use functions to get the data. 
The right to run functions on a particular time series is linked to access levels. 

### Access Levels

{timeseriesdb} can control access at the time series and even at the version level. 
The basic installation ships with three access levels: public, main and restricted. 
The idea of the public level is that every users with access to the {timeseriesdb}
database schema can read public series. The idea of the 'main' level is to give 
all users inside an organization access to the assigned data. 'restricted' is
designed to limit access to subgroups within an organization. We guess you see 
where this is going. By adding further access levels, admins can add more narrow
user groups for parts of the data.

[db_access_level_list()](../reference/db_access_level_list.html) lists all access levels available for a schema. 
New levels are created easily. Note also that new levels can be made the default.

```{r,eval=FALSE}
db_access_level_create(con, 
                       access_level_name =  "management",
                       access_level_description = "series only available",
                       access_level_default = TRUE,
                       schema = "tsdb_test"
                       )
```

Obviously only one access level can be 'default' at a time. Defaults can be
set post creation as well using [db_access_level_set_default()](../reference/db_access_level_set_default.html). To delete levels, 
run [db_access_level_delete()](../reference/db_access_level_delete.html).

### Assigning Access

To assign an access level to a single time series use [db_ts_change_access()](../reference/db_ts_change_access.html), 
to assign an access level to an entire dataset, i.e., to all series in a set, use
[db_dataset_change_access()](../reference/db_dataset_change_access.html).









