---
title: "Advanced Usage"
author: "Matthias Bannert and Severin Thöni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Starting Point

We assume you are past the basics and R, {timeseriesdb} and PostgreSQL are up and running. 
(*For the remainder of this guide we assume PostgreSQL runs on a local docker container accessible via port 1111.*)

## Datasets

In {timeseriesdb}, all time series have to be part of a dataset. If the user 
stores a time series w/o specifying a dataset that series is assigned to the 'default'
dataset. Datasets allow to group data and store data descriptions at the 
dataset level. Also, datasets come in handy for admin operations like 
deletion of an entire dataset. 


Let's create a new dataset!

```{r, eval=FALSE}
library(kofdata)
arr <- get_time_series("ch.zrh_airport.arrival.total")
dep <- get_time_series("ch.zrh_airport.departure.total")

store_time_series(con, c(arr,dep), schema = "tsdb_test")

db_create_dataset(con, "ch.zrh_airport",
                  set_description = "Arrivals and Departures ZH Airport",
                  schema = "tsdb_test")

# see, the new dataset is there... 
db_list_datasets(con, schema = "tsdb_test")

```
Apparently, the dataset is empty. We need to assign some time series first.
Make sure the time series you want to register to a set are actually around.
Unlike `store_time_series` assigning series to a dataset is *not* a
vintage level operation, it applies to all versions of a series. This is why time series are assigned to the 'default' schema at first when they are stored and can be moved later on. 

```{r, eval=FALSE}
db_assign_dataset(con,
                  ts_keys = c("ch.zrh_airport.arrival.total",
                              "ch.zrh_airport.departure.total"),
                  set_name = "ch.zrh_airport",
                  schema = "tsdb_test")

```
```{r}
db_get_dataset_keys(con, "ch.zrh_airport", schema = "tsdb_test")
```

```{r}
db_get_dataset_id(con, "ch.zrh_airport.arrival.total",
                  schema = "tsdb_test")
```


## Release Management







## Collections

Collections are user specific compositions of time series identifiers. Users 
of earlier versions of time series may think of this functionality as 'sets'. 
To avoid any confusion, let's be clear about the new terminology here: 

- datasets = registration of time series into global datasets at the admin level. 
time series are unique ACROSS datasets.

- collection = playlist like user specific bookmarks. A time series can part of 
multiple collections, users can have multiple collections. 


## Advanced Meta Information

### Multi Language Data Descriptions


### Levels of Assignment


## Versioning of Time Series (Vintages)




## Access Levels and Rights Management


## Hygenie and Maintenance






## Remove 




{timeseriesdb} keeps track of different versions of a time series. This feature was designed with regular releases of data in mind. In official and economic statistics, releases do not simply mean to append another observation at end of time series. Seasonal adjustment and other filters as well data corrections cause revisions of a time series. In order to keep research, e.g., forecasts reproducible, it is important to archive all versions of a time series. Doing so, allows to assess forecast performance and makes sure the state of information at the time the forecast was made can be reproduced.

{timeseriesdb} allows to store one version of a time series per day, i.e., series 



## Collections (aka playlists)
## Datasets

In {timeseriesdb} *every* time series belongs to a dataset. If no dataset is specified when a time series is stored, the series automagically gets assigned to the *default* dataset.

The idea behind datasets is to group time series according to their source and allow for group level data description. The
following helpers help you to explore which keys are in a given set and which keys are in which set.

```{r, eval=FALSE}
db_get_dataset_keys(con, set_name = "default",
                    schema = "tsdb_test")
db_get_dataset_id(con, c("AirPassengers", "JohnsonJohnson"),
                  schema = "tsdb_test")
```

You can easily add new datasets. Note though, that dataset are not user specific, but global groups. The user specific analog to sets is called collection in will be described later on. 

```{r, eval=FALSE}
db_create_dataset(con, "stocks",
                  set_description = "Stock market returns.",
                  schema = "tsdb_test")
```
Once a set is created you can move existing series to that set. 

```{r, eval=FALSE}
db_assign_dataset(con,"JohnsonJohnson","stocks",
                  schema = "tsdb_test")
db_get_dataset_id(con, c("AirPassengers", "JohnsonJohnson"), schema = "tsdb_test")
```
