---
title: "Basic Usage"
subtitle: "Getting Started}"
author: "Matthias Bannert and Severin Thöni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Starting Point

This guide assumes you have installed the {timeseriesdb} R package and a local PostgreSQL client library successfully. We also assume that you have a PostgreSQL database running somewhere that you can connect to. (*For the remainder of this guide we assume PostgreSQL runs on a local docker container accessible via port 1111.*)


## Store Time Series 

Let's take two time series from the {datasets} R package that ships with R (monthly AirPassengers, JohnsonJohnson quarterly earnings per share) and put them into a named list. The {timeseriesdb} R package uses lists of time series or long format time series `data.tables` to store one or more time series into PostgreSQL. 

```{r, eval=FALSE}
library(timeseriesdb)

con <- dbConnect(RPostgres::Postgres(),
                 dbname = "postgres",
                 user = "dev_admin",
                 host = "localhost",
                 password = "dev_admin",
                 port = 1111)
tsl <- list(AirPassengers = AirPassengers,
            JohnsonJohnson = JohnsonJohnson)
store_time_series(con, tsl, "main",
                  schema = "tsdb_test")

```
 
## Read Time Series
 
Now let's read the series we have just stored to the database into R again. 
Note that `read_time_series` returns a list of time series whose elements' names
are the time series keys (unique identifiers) from the database. In this case
example we use our [{tstools}]() package to visualize both series. 
 
```{r, eval=FALSE}

out <- read_time_series(con, c("JohnsonJohnson",
                               "AirPassengers"),
                        schema = "tsdb_test")

library(tstools)
tsplot(AirPassengers = out$AirPassengers,
       tsr = list(JohnsonJohnson = out$JohnsonJohnson))
```

```{r,echo=FALSE}
library(tstools)
out <- list(AirPassengers = AirPassengers,
            JohnsonJohnson = JohnsonJohnson)
tsplot(AirPassengers = out$AirPassengers,
       tsr = list(JohnsonJohnson = out$JohnsonJohnson))



```



## Basic Metadata 

To most basic way to store data descriptions is to assign meta information 
at the time series level. 

```{r, eval=FALSE}
db_read_ts_metadata(con, "AirPassengers",
                    schema = "tsdb_test")
```


## Advanced Use

{timeseriesdb} offers a plethora of features beyond mere storage of time series and 
their descriptions: {timeseriesdb} can keep track of different versions of the 
same time series (vintages), group time series into datasets, store user specific 
collection of series, manage access on time series level and manage release dates. 

The learn more about these advanced features read the [advanced usage vignette]().








